---
title: "label_transfer_others"
output: html_document
---
Further label transfers
```{r}
library(Seurat)
library(tidyverse)
library(data.table)
library(patchwork)
library(gridExtra)
library(grid)
library(openxlsx)

id2ensembl <- fread('/home/cbmr/qwn903/ygg-projects/amj/tune/timshel-bmicelltypes/data/gene_annotations/Mus_musculus.GRCm38.90.gene_name_version2ensembl.txt.gz')
```

### Datatset edits
Mikkelsen
```{r}
mik <- readRDS("/data/pub-others/mikkelsen-natureneuroscience-2019/output/mikkelsen_natureneuroscience_2019_seuratObj.RDS.gz")
mik_meta <- fread("/data/pub-others/mikkelsen-natureneuroscience-2019/mikkelsen_natureneuroscience_2019_1_neurons_metadata.csv.gz")
mik_meta %>% column_to_rownames("Cell.ID") -> mik_meta
```
Edit Mikkelsen dataset to have ensembl ids
Subset to only cells with metadata
```{r}
mik.dat <- as.data.frame(as.matrix(mik@assays$RNA@counts)) %>% rownames_to_column("gene_name_optimal")
id_join <- left_join(mik.dat, id2ensembl[, c(1,2)])
mik.dat <- id_join[, c(7216, 2:7215)] %>% na.omit %>% remove_rownames() %>% column_to_rownames('ensembl_gene_id')

mik <- CreateSeuratObject(counts = mik.dat, meta.data = mik_meta)
mik_sub <- subset(mik, cell_type %in% mik_meta$cell_type)

mik_sub <- SCTransform(mik_sub)

# Save object
save(mik_sub, file = "./dat/mikkelsen_seurat_obj.RData")
```

Moffett
```{r}
mof.dat <- fread("/data/pub-others/moffitt-science-2018/moffitt_all_cells_raw_ensembl.csv.gz", header = T, sep = ",") %>% column_to_rownames('ensembl')
mof_meta <- fread("/data/pub-others/moffitt-science-2018/moffitt-science-2018_1_metadata.csv.gz")
mof_meta %>% column_to_rownames("Cell.name") -> mof_meta

mof <- CreateSeuratObject(counts = mof.dat, meta.data = mof_meta)
mof <- SCTransform(mof)
# Save object
save(mof, file = "./dat/mof_seurat_obj.RData")
```

Kim 10x
```{r}
kim.dat <- fread("~/ygg-projects/amj/my_projects/kimVMH2019/processed/kimVMH2019_10x.umi.csv.gz")
kim_meta <- fread("~/ygg-projects/amj/my_projects/kimVMH2019/processed/kimVMH2019_10x.metadata.csv")
kim_meta %>% column_to_rownames("cell_id") -> kim_meta
```
Edit Kim dataset to have ensembl ids
```{r}
colnames(kim.dat)[1] <- "gene_name_optimal"
id_join <- left_join(kim.dat, id2ensembl[, c(1,2)])
kim.dat <- id_join[,c(41387,2:41386)] %>% na.omit %>% remove_rownames() %>% column_to_rownames('ensembl_gene_id')
kim <- CreateSeuratObject(counts = kim.dat, meta.data = kim_meta)
kim <- SCTransform(kim)
# Save object
save(kim, file = "./dat/kim10x_seurat_obj.RData")
```

Kim smartseq
```{r}
kim.dat <- fread("~/ygg-projects/amj/my_projects/kimVMH2019/processed/kimVMH2019_smartseq.umi.csv.gz")
kim_meta <- fread("~/ygg-projects/amj/my_projects/kimVMH2019/processed/kimVMH2019_smartseq.metadata.csv")
kim_meta %>% column_to_rownames("cell_id") -> kim_meta
```
Edit Kim dataset to have ensembl ids
```{r}
colnames(kim.dat)[1] <- "gene_name_optimal"
id_join <- left_join(kim.dat, id2ensembl[, c(1,2)])
kim.dat <- id_join[,c(4576,2:4575)] %>% na.omit %>% remove_rownames() %>% column_to_rownames('ensembl_gene_id')
kim <- CreateSeuratObject(counts = kim.dat, meta.data = kim_meta)
kim <- SCTransform(kim)
# Save object
save(kim, file = "./dat/kimsmartseq_seurat_obj.RData")
```

### Label transfer
```{r}
load('./dat/tomcells_r_adan_seurat_obj.RData')
# Remove cells with no cell-calling from R. Adan
seur_obj <- subset(seur_obj, cells = na.omit(seur_obj@meta.data$V1))
```
Run together
```{r}
# Mikkelsen
load('./dat/mikkelsen_seurat_obj.RData')
hyp.anchors <- FindTransferAnchors(reference = mik_sub, normalization.method = "SCT",
                                   query = seur_obj, dims = 1:30, reference.assay = "SCT", 
                                   query.assay = "SCT", reduction = "cca")
predictions_mik <- TransferData(anchorset = hyp.anchors, 
                                refdata = mik_sub$cell_type, dims = 1:30, weight.reduction = "cca")

# Moffett
load('./dat/mof_seurat_obj.RData')
hyp.anchors <- FindTransferAnchors(reference = mof, normalization.method = "SCT",
                                   query = seur_obj, dims = 1:30, reference.assay = "SCT", 
                                   query.assay = "SCT", reduction = "cca")
predictions_mof <- TransferData(anchorset = hyp.anchors, 
                                refdata = mof$cell_type, dims = 1:30, weight.reduction = "cca")

# Kim smartseq
load('./dat/kimsmartseq_seurat_obj.RData')
hyp.anchors <- FindTransferAnchors(reference = kim, normalization.method = "SCT",
                                   query = seur_obj, dims = 1:30, reference.assay = "SCT", 
                                   query.assay = "SCT", reduction = "cca")
predictions_kim <- TransferData(anchorset = hyp.anchors, 
                                refdata = kim$cell_type, dims = 1:30, weight.reduction = "cca")

# Kim 10x
load('./dat/kim10x_seurat_obj.RData')
hyp.anchors <- FindTransferAnchors(reference = kim, normalization.method = "SCT",
                                   query = seur_obj, dims = 1:30, reference.assay = "SCT", 
                                   query.assay = "SCT", reduction = "cca")
predictions_kim10x <- TransferData(anchorset = hyp.anchors, 
                                refdata = kim$cell_type, dims = 1:30, weight.reduction = "cca")


# Editing predictions
predictions_mik %>% rownames_to_column("V1") %>%
  mutate(max_pred_mik = ifelse(prediction.score.max > 0.5, predicted.id, NA),
         pred_score_mik = ifelse(prediction.score.max > 0.5, prediction.score.max, NA)) %>%
  select(V1, max_pred_mik, pred_score_mik) -> pred1

predictions_mof %>% rownames_to_column("V1")%>% filter(prediction.score.max > 0.5) %>% 
  rename(max_pred_mof = predicted.id,
         pred_score_mof = prediction.score.max) %>%
  select(V1, max_pred_mof, pred_score_mof) -> pred2

predictions_kim %>% rownames_to_column("V1") %>% filter(prediction.score.max > 0.5) %>% 
  rename(max_pred_kim = predicted.id,
         pred_score_kim = prediction.score.max) %>%
  select(V1, max_pred_kim, pred_score_kim) -> pred3

predictions_kim10x %>% rownames_to_column("V1") %>% filter(prediction.score.max > 0.5) %>% 
  rename(max_pred_kim10 = predicted.id,
         pred_score_kim10 = prediction.score.max) %>%
  select(V1, max_pred_kim10, pred_score_kim10) -> pred4


predictions_all <- left_join(pred1, pred2) %>% left_join(., pred3) %>%
  left_join(., pred4) %>% column_to_rownames('V1')

```
Adding predictions as metadata
```{r}
load('./dat/tomcells_w.labels_r_adan_seurat_obj.RData')
seur_obj@meta.data
seur_obj <- AddMetaData(seur_obj, metadata = predictions_all)

# Save object
save(seur_obj, file = "./dat/tomcells_w.labels_extra_r_adan_seurat_obj.RData")
```

### Visualization
All cells
```{r}
load("./dat/tomcells_w.labels_extra_r_adan_seurat_obj.RData")
seur_obj <- RunPCA(seur_obj, verbose = F)
seur_obj <- RunUMAP(seur_obj, dims = 1:30)

seur_obj <- FindNeighbors(seur_obj, dims = 1:30)
seur_obj <- FindClusters(seur_obj, resolution = 2.5)


p1 <- DimPlot(seur_obj, label = TRUE, group.by = 'cell_types', label.size = 6, pt.size = 3, repel = T) + 
  theme(legend.position = "bottom") + ggtitle("R Adan celltypes")

p2 <- DimPlot(seur_obj, label = T, group.by = 'max_pred_mik', label.size = 6, pt.size = 3, repel = T) + 
  theme(legend.position = "bottom") + ggtitle("Mikkelsen labels")

p3 <- DimPlot(seur_obj, label = T, group.by = 'max_pred_mof', label.size = 6, pt.size = 3, repel = T) + 
  theme(legend.position = "bottom")+ ggtitle("Moffett labels")

p4 <- DimPlot(seur_obj, label = T, group.by = 'max_pred_kim', label.size = 6, pt.size = 3, repel = T) + 
  theme(legend.position = "bottom") + ggtitle("Kim smartseq labels")

p5 <- DimPlot(seur_obj, label = T, group.by = 'max_pred_kim10', label.size = 6, pt.size = 3, repel = T) + 
  theme(legend.position = "bottom") + ggtitle("Kim 10x labels")
```

Only neurons
```{r}
# Neuronal clusters only
seur_obj_neuro <- subset(seur_obj, cells = seur_obj@meta.data[grepl('neurons', seur_obj$cell_types), 'V1'])
# These are now standard steps in the Seurat workflow for visualization and clustering
seur_obj_neuro <- RunPCA(seur_obj_neuro, verbose = F)
seur_obj_neuro <- RunUMAP(seur_obj_neuro, dims = 1:30)
seur_obj_neuro <- FindNeighbors(seur_obj_neuro, dims = 1:30)
seur_obj_neuro <- FindClusters(seur_obj_neuro, resolution = 1.7)

p_n1 <- DimPlot(seur_obj_neuro, group.by = "cell_types", label = T, label.size = 4, pt.size = 3, repel = T) + 
  theme(legend.position = "bottom") 

p_n2 <- DimPlot(seur_obj_neuro, group.by = "max_pred_mik", label = T, label.size = 4, pt.size = 3, repel = T) + 
  theme(legend.position = "bottom") + ggtitle("Mikkelsen")

p_n3 <- DimPlot(seur_obj_neuro, group.by = "max_pred_mof", label = T, label.size = 4, pt.size = 3, repel = T) + 
  theme(legend.position = "bottom") + ggtitle("Moffett")

p_n4 <- DimPlot(seur_obj_neuro, group.by = "max_pred_kim", label = T, label.size = 4, pt.size = 3, repel = T) + 
  theme(legend.position = "bottom") + ggtitle("Kim smartseq")

p_n5 <- DimPlot(seur_obj_neuro, group.by = "max_pred_kim10", label = T, label.size = 4, pt.size = 3, repel = T) + theme(legend.position = "bottom")+ ggtitle("Kim 10x")
```


Writing output
```{r}
pdf('./output/Label_transfer_all.pdf', width = 17, height = 10)
# R_adan celltypes
p1 + p2 + plot_annotation(
  title = 'Labeltransfer',
  subtitle = 'Comparing labels from R. Adan with Mikkelsen labels - max.pred 50%')
p1 + p3 + plot_annotation(
  title = 'Labeltransfer',
  subtitle = 'Comparing labels from R. Adan with Moffett labels - max.pred 50%')
p1 + p4 + plot_annotation(
  title = 'Labeltransfer',
  subtitle = 'Comparing labels from R. Adan with Kim smartseq labels - max.pred 50%')
p1 + p5 + plot_annotation(
  title = 'Labeltransfer',
  subtitle = 'Comparing labels from R. Adan with Kim 10x labels - max.pred 50%')
p_n1 + p_n2 + plot_annotation(
  title = 'Labeltransfer, Neuronal clusters only',
  subtitle = 'Comparing labels from R. Adan with Mikkelsen labels - max.pred 50%, neuronal subcluster')
p_n1 + p_n3 + plot_annotation(
  title = 'Labeltransfer, Neuronal clusters only',
  subtitle = 'Comparing labels from R. Adan with Moffett labels - max.pred 50%, neuronal subcluster')
p_n1 + p_n4 + plot_annotation(
  title = 'Labeltransfer, Neuronal clusters only',
  subtitle = 'Comparing labels from R. Adan with Kim smartseq labels - max.pred 50%, neuronal subcluster')
p_n1 + p_n5 + plot_annotation(
  title = 'Labeltransfer, Neuronal clusters only',
  subtitle = 'Comparing labels from R. Adan with Kim 10x labels - max.pred 50%, neuronal subcluster')
dev.off()
```

Highlighting interesting mappings.
Only include clusters in which over 20% of the cells have prediction-scores above 75%
```{r}
h_tab <- seur_obj@meta.data

h_tab %>% add_count(cell_types) %>% group_by(cell_types) %>%
  mutate_at(vars(starts_with('pred_score')), function(x)sum(as.numeric(x)>0.75, na.rm = T)) %>% 
  mutate(r_mik = pred_score_mik/n,
         r_mof = pred_score_mof/n,
         r_kim = pred_score_kim/n,
         r_kim10 = pred_score_kim10/n,) %>% 
  select(starts_with('r_')) %>% mutate_at(vars(starts_with("r_")), function(x)x>0.2) %>% unique -> h_out

h_tab[which(h_tab$cell_types %in% h_out[which(h_out$r_kim>0.25),]$cell_types),]
```

### Tables
```{r}
tab <- seur_obj@meta.data
tab_mik <- tab %>% group_by(cell_types) %>% count(max_pred_mik) %>% filter(n>1) %>% arrange(cell_types, desc(n)) %>%
  left_join(., h_out[,c(1,2)])

tab_mof <- tab %>% group_by(cell_types) %>% count(max_pred_mof) %>% filter(n>1) %>% arrange(cell_types, desc(n))%>%
  left_join(., h_out[,c(1,3)])

tab_kim <- tab %>% group_by(cell_types) %>% count(max_pred_kim) %>% filter(n>1) %>% arrange(cell_types, desc(n))%>%
  left_join(., h_out[,c(1,4)])

tab_kim10 <- tab %>% group_by(cell_types) %>% count(max_pred_kim10) %>% filter(n>1) %>% arrange(cell_types, desc(n))%>%
  left_join(., h_out[,c(1,5)])

tab_mik$cell_types[duplicated(tab_mik$cell_types)] <- ""
tab_mof$cell_types[duplicated(tab_mof$cell_types)] <- ""
tab_kim$cell_types[duplicated(tab_kim$cell_types)] <- ""
tab_kim10$cell_types[duplicated(tab_kim10$cell_types)] <- ""


pdf('./output/Label_transfer_highlights.pdf')
grid.arrange(textGrob("Mikkelsen labeltransfer", just = "top", gp=gpar(fontsize = 14)), 
             tableGrob(tab_mik[grepl(TRUE, tab_mik$r_mik),1:3], rows = NULL), heights = c(0.05,1))
grid.arrange(textGrob("Moffett labeltransfer", just = "top", gp=gpar(fontsize = 14)), 
             tableGrob(tab_mof[grepl(TRUE, tab_mof$r_mof),1:3], rows = NULL), heights = c(0.05,1))
grid.arrange(textGrob("Kim smartseq labeltransfer", just = "top", gp=gpar(fontsize = 14)), 
             tableGrob(tab_kim[grepl(TRUE, tab_kim$r_kim),1:3], rows = NULL), heights = c(0.05,1))
# grid.arrange(textGrob("Kim 10x labeltransfer", just = "top", gp=gpar(fontsize = 14)),
#              tableGrob(tab_kim10[grepl(TRUE, tab_kim10$r_kim10),1:3], rows = NULL), heights = c(0.05,1))
# No cells from Kim10x in the highlights table output
dev.off()

```

Write excel with all mapping information
```{r}
# Make a column with the mapping prediction score
tab <- seur_obj@meta.data

tab %>% select('cell_types', starts_with('max_pred'), starts_with('pred_score')) %>%
  select(-"pred_score_neuron1.5", -"max_pred_neuron1.5" )-> all_tabs

write.xlsx(all_tabs, file = './output/all_extra_labeltransfer.xlsx')

```

